---
title: "Los Angeles Education"
author: "Inam Khan"
execute: 
  echo: false
---

```{r}
#| message: false

library(tidycensus)
library(tidyverse)
library(leaflet)
library(sf)
library(viridis)
```

```{r}
#| warning: false

x <- get_acs(
  geography = "tract",
  variables = c(total = "B15003_001", 
                bachelor = "B15003_022", 
                masters = "B15003_023",
                professional = "B15003_024", 
                doctorate = "B15003_025"),
  county = "Los Angeles",
  state = "CA",
  year = 2022,
  geometry = TRUE,
  progress_bar = FALSE, 
  output = "wide"
) |> 
  mutate(pct_bachelors_or_higher = 100 * (
    bachelorE + mastersE + professionalE + doctorateE) / totalE)

write_rds(x, "la.rds")
```

```{r}
#| warning: false

# Create color palette based on bachelors percentage
pal <- colorNumeric(
  palette = "viridis",
  domain = x$pct_bachelors_or_higher,
  na.color = "#808080"
)

# Create labels for popups
labels <- sprintf(
  "<strong>Tract %s</strong><br/>
  Total population 25+: %s<br/>
  Bachelor's degrees: %s (%.1f%%)<br/>
  Master's degrees: %s (%.1f%%)<br/>
  Professional degrees: %s (%.1f%%)<br/>
  Doctorate degrees: %s (%.1f%%)<br/>
  <strong>Bachelor's or higher: %.1f%%</strong>",
  substr(x$GEOID, 6, 11),  # Just show the tract number
  format(x$totalE, big.mark = ","),
  format(x$bachelorE, big.mark = ","), round(100 * x$bachelorE/x$totalE, 1),
  format(x$mastersE, big.mark = ","), round(100 * x$mastersE/x$totalE, 1),
  format(x$professionalE, big.mark = ","), round(100 * x$professionalE/x$totalE, 1),
  format(x$doctorateE, big.mark = ","), round(100 * x$doctorateE/x$totalE, 1),
  x$pct_bachelors_or_higher
) %>% lapply(htmltools::HTML)

# Create the map
leaflet(x) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    fillColor = ~pal(pct_bachelors_or_higher),
    weight = 0.5,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    highlight = highlightOptions(
      weight = 2,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.9,
      bringToFront = TRUE
    ),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto"
    )
  ) %>%
  addLegend(
    pal = pal,
    values = ~pct_bachelors_or_higher,
    opacity = 0.7,
    title = "% with Bachelor's<br/>or Higher",
    position = "bottomright",
    labFormat = labelFormat(suffix = "%")
  ) %>%
  setView(lng = -118.2437, lat = 34.0522, zoom = 9) %>%
  addControl(
    "<strong>Educational Attainment in Los Angeles</strong><br/>Percentage of population 25+ with bachelor's degree or higher by census tract",
    position = "topright"
  )
```

This visualization presents educational attainment across Los Angeles census tracts, focusing on the percentage of the population aged 25 and older with a bachelor's degree or higher. Created by Inam Khan, the map highlights significant geographic disparities in educational achievement throughout the region. The legend indicates percentages ranging from 10% to 90%, allowing viewers to compare attainment levels across different neighborhoods. Notable areas like Irvine, Mission Viejo, and Thousand Oaks appear to have higher educational attainment, while other regions show more modest percentages. The map also includes demographic references to White, Hispanic, and other population groups, suggesting potential correlations between education levels and community composition. Built using Leaflet with OpenStreetMap and CARTO basemaps, this interactive tool provides valuable insights into the spatial distribution of educational achievement in one of America's most populous metropolitan areas. The inclusion of specific city names like Santa Ana, Anaheim, and Huntington Beach helps orient users while examining the educational landscape of Southern California.

